### 2022-06-20 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@85ea305586fa11eef39b9fb32a52100fd76a1565 🚀

Previous studies that have used host richness as proxy have predicted a
high diversity of unsampled bat viruses {[}Olival{]}, bat coronaviruses
{[}Anthony{]}, and even specifically betacoronaviruses {[}Becker{]} in
both the Amazon and southeast Asia. While we find that both regions are
characterized by highly divergent host and viral communities, our
framework identifies key differences between the regions. We find that
Latin America is a hotspot of both host and viral distinctiveness,
suggesting that this branch of the bat-betacoronavirus complex may be
undergoing independent evolutionary dynamics from the rest of the global
pool, but with limited potential for viral diversification--- a finding
that is supported by previous work indicating a higher rate of
codivergence in Latin America {[}Anthony{]}. In contrast, in southeast
Asia, host richness and viral distinctiveness are high but sharing is
low; this suggests a different type of evolutionary dynamics that could
generate high viral diversity locally (see \emph{e.g.} (Latinne et al.
2020), as well as the discovery of recombinant viruses that share
genetic material from both the SARS-CoV and SARS-CoV-2 branches of the
Sarbecovirus lineage {[}Wu{]}). Both of these regions are priority areas
for sampling, especially given predictions by Becker et al. (2022) that
they contain many bat hosts of undiscovered betacoronaviruses. However,
both the evolutionary and ecological aspects of emergence risk are
likely higher in southeast Asia---a fact that will only become more
relevant, as bats track shifting climates and exchange viruses with
other species, creating a hotspot of cross-species transmission unique
to the region {[}Carlson{]}.

{[}para about conservation{]} There are several factors that drive
changes in the diversity of bats (Alves et al. 2018), but human
activities' effects on the ecosystem (like modifications of land use)
could significantly decrease it. Therefore, it can be suggested that
changes in the diversity of betacoronaviruses in bats are linked to
their biogeographic variation, and human population density and other
anthropogenic factors are decisive moderators for its implications in
public health. With the increase of contact between humans and potential
hosts, we also increase the risk of emergence of novel diseases (Johnson
et al. 2020), as previous studies on RNA viruses suggest the importance
of host phylogeography at the time of virus dispersal (Gryseels et al.
2017). One of these scenarios where interaction between bats and humans
can occur can be seed dispersal in tropical agroecosystems. It opens the
discussion of whether the fruits thrown by bats not only disperse seeds
but could also be a source of indirect interaction between viruses of
bat origin and humans (Deshpande et al. 2022). This represents a
challenge for conservation strategies and disease ecology since some
areas can haveboth potential for the acquisition of zoonotic viruses and
bat-human interactions; in particular, the challenge lies in the fact
that actual exposure must then be quantified accounting for several
transmission scenarios, including both direct and indirect bat - human
interaction.

@@

We downloaded the rangemap of every current bat species that was
classified as an empirically documented host of \emph{Betacoronavirus}
from the previous step, according to recent IUCN data (IUCN 2021). The
range maps were subsequently rasterized using the \texttt{rasterize}
function from \texttt{GDAL} (Rouault et al. 2022) at a resolution of
approximately 100kmx100km. For every pixel in the resulting raster where
at least one bat host of \emph{Betacoronavirus} was present, we extract
the species pool (list of all competent bat hosts), which was used to
calculate the following risk assessment components: bat phylogenetic
diversity, bat compositional uniqueness, and predicted viral sharing
risk.

>>>>

Here, we propose a simple freamework with broad explanatory power that
helps contextualize discoveries like highly divergent nobecoviruses in
Madagascar and the previously-neglected adaptive radiation of
sarbecoviruses outside of southern China and throughout southeast Asia.
In doing so, it advances ecological theory beyond the current state of
the art for global maps of emergence risk. For example, previous studies
that have used host richness as proxy have predicted a high diversity of
unsampled bat viruses {[}Olival{]}, bat coronaviruses {[}Anthony{]}, and
even specifically betacoronaviruses {[}Becker{]} in both the Amazon and
southeast Asia. While we find that both regions are characterized by
highly divergent host and viral communities, our framework identifies
key differences between the regions. We find that Latin America is a
hotspot of both host and viral distinctiveness, suggesting that this
branch of the bat-betacoronavirus complex may be undergoing independent
evolutionary dynamics from the rest of the global pool, but with limited
potential for viral diversification--- a finding that is supported by
previous work indicating a higher rate of codivergence in Latin America
{[}Anthony{]}. In contrast, in southeast Asia, host richness and viral
distinctiveness are high but sharing is low; this suggests a different
type of evolutionary dynamics that could generate high local diversity
of viruses through host switching and viral recombination (see
\emph{e.g.} (Latinne et al. 2020), as well as the discovery of
recombinant viruses that share genetic material from both the SARS-CoV
and SARS-CoV-2 branches of the Sarbecovirus lineage {[}Wu{]}). Both of
these regions are priority areas for sampling, especially given
predictions by Becker et al. (2022) that they contain many bat hosts of
undiscovered betacoronaviruses. However, both the evolutionary and
ecological aspects of emergence risk are likely higher in southeast
Asia---a fact that will only become more relevant, as bats track
shifting climates and exchange viruses with other species, creating a
hotspot of cross-species transmission unique to the region
{[}Carlson{]}.

{[}Tim tinker with para about conservation{]} There are several factors
that drive changes in the diversity of bats (Alves et al. 2018), but
human activities' effects on the ecosystem (like modifications of land
use) could significantly decrease it. Therefore, it can be suggested
that changes in the diversity of betacoronaviruses in bats are linked to
their biogeographic variation, and human population density and other
anthropogenic factors are decisive moderators for its implications in
public health. With the increase of contact between humans and potential
hosts, we also increase the risk of emergence of novel diseases (Johnson
et al. 2020), as previous studies on RNA viruses suggest the importance
of host phylogeography at the time of virus dispersal (Gryseels et al.
2017). One of these scenarios where interaction between bats and humans
can occur can be seed dispersal in tropical agroecosystems. It opens the
discussion of whether the fruits thrown by bats not only disperse seeds
but could also be a source of indirect interaction between viruses of
bat origin and humans (Deshpande et al. 2022). This represents a
challenge for conservation strategies and disease ecology since some
areas can haveboth potential for the acquisition of zoonotic viruses and
bat-human interactions; in particular, the challenge lies in the fact
that actual exposure must then be quantified accounting for several
transmission scenarios, including both direct and indirect bat - human
interaction.

@@

We downloaded the rangemap of every current bat species that was
classified as an empirically documented host of \emph{Betacoronavirus}
from the previous step, according to recent IUCN data (IUCN 2021). The
range maps were subsequently rasterized using the \texttt{rasterize}
function from \texttt{GDAL} (Rouault et al. 2022) at a resolution of
approximately 100kmx100km. For every pixel in the resulting raster where
at least one bat host of \emph{Betacoronavirus} was present, we extract
the species pool (list of all known bat hosts), which was used to
calculate the following risk assessment components: bat phylogenetic
diversity, bat compositional uniqueness, and predicted viral sharing
risk.
